\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-09-06 smurf_paper.tex
Add a bit of gravitas to the Starlink citation
explicitly mention kapuahi release
No reference to it as such since that would only be
(Berry et al, 2013, in preparation)
Use hyperref and add some URLs. Also add a final reference

@@

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{holland2012} is a new instrument for the 15-m James
Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
simultaneously images the sky in two broad bands centered over 450 and
850\,\micron, with approximately 7.5 and 14.5\,arcsec full-width at
half-maximum (FWHM) point spread functions (PSFs). The focal planes at
each wavelength are populated with 4 rectangular subarrays, consisting
of $40 \times 32$ bolometers each, and together subtend a nearly
7\,arcmin $\times$ 7\,arcmin footprint on the sky (excluding gaps, the
continuous solid angle is about 43\,arcmin$^{2}$ per focal plane).
This paper describes the properties of \scuba\ data that are relevant
to producing maps of the imaging data, and the Submillimetre User
Reduction Facility, SMURF, a software package for performing the
reduction written using the Starlink Software Environment
\citep{2009ASPC..411..418J}. The details of the instrument design,
performance, and calibration are given in two companion papers:
\citet{holland2012} and \citet{dempsey2012}.

@@

This paper is organized in the following way. We first describe the
properties of \scuba\ data in Section~\ref{sec:data} to motivate the
data model that we have adopted. Next, the details of the SMURF
algorithm are given in Section~\ref{sec:algorithm}. The paper is
concluded in Section~\ref{sec:examples} with three detailed test cases
that span the majority of observation types likely to be undertaken
with \scuba, with an emphasis on the mitigation of divergence
problems, and characterizing the output maps: (i) Uranus, a bright,
compact source (Section~\ref{sec:point}); (ii) the Lockman Hole, a
blind survey of faint point-like source (Section~\ref{sec:cosmo}); and
(iii) the star-forming region M17, including bright, extended emission
(Section~\ref{sec:extended}). All of the data analysed in this paper
are publicly available through the CADC \scuba\ raw-data queries page
for the dates and observation numbers given in the
text\footnote{\texttt{http://www3.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/
jcmt/search/scuba2}}. All of the analysis was performed using
\texttt{Starlink} version ... .

@@

As described in the introduction, the approach taken by SMURF to
reducing \scuba\ data is to model (and remove) predominantly
low-frequency noise sources that are correlated amongst detectors, and
to iterate this process along with estimates of the map. Significant
experimentation with different models for noise sources during \scuba\
commissioning required a highly flexible software framework, and a
configurable user interface. To achieve this goal, while minimizing
development time, it was decided to build SMURF as a Starlink package
\citep{2009ASPC..411..418J}, which provides access to a large suite of
libraries (including a commanding and messaging interface, astrometric
coordinate conversions and generation of standard WCS information,
file formats etc.). Furthermore, Starlink is open source, and already
used extensively at the Joint Astronomy Centre (host of the JCMT) for
many other systems, which helps with interoperability. Though
originally written in FORTRAN, many of the core Starlink libraries are
now ported to native C, or at least have a C interface. It was
therefore decided to develop SMURF in C as well (rather than C++, for
example, which would have required adding further dependencies to
Starlink), although we have taken an object-oriented approach. For
example, all of the data for a given noise model is encapsulated in a
C structure, and there is a standard interface for all functions that
handle models (member data and functions for the class,
respectively). In this way it is easy to extend SMURF with new
models. Parallelisation is incorporated in the most time-consuming
low-level routines using threads (e.g., performing Fast Fourier
Transforms, re-gridding the data), usually handling either independent
blocks of bolometers, or blocks of time, in each thread, depending on
the nature of the calculation. \textbf{... something about scalability
...}

>>>>

%% Recommended by MN for handling hyperlinks
\usepackage[pdftex,pdfpagemode={UseOutlines},bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}

@@

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{holland2012} is a new instrument for the 15-m James
Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
simultaneously images the sky in two broad bands centered over 450 and
850\,\micron, with approximately 7.5 and 14.5\,arcsec full-width at
half-maximum (FWHM) point spread functions (PSFs). The focal planes at
each wavelength are populated with 4 rectangular subarrays, consisting
of $40 \times 32$ bolometers each, and together subtend a nearly
7\,arcmin $\times$ 7\,arcmin footprint on the sky (excluding gaps, the
continuous solid angle is about 43\,arcmin$^{2}$ per focal plane).
This paper describes the properties of \scuba\ data that are relevant
to producing maps of the imaging data, and the Submillimetre User
Reduction Facility, SMURF, a software package for performing the
reduction written using the Starlink Software Environment
\citep{1993ASPC...52..229W,2009ASPC..411..418J}. The details of the instrument design,
performance, and calibration are given in two companion papers:
\citet{holland2012} and \citet{dempsey2012}.

@@

This paper is organized in the following way. We first describe the
properties of \scuba\ data in Section~\ref{sec:data} to motivate the
data model that we have adopted. Next, the details of the SMURF
algorithm are given in Section~\ref{sec:algorithm}. The paper is
concluded in Section~\ref{sec:examples} with three detailed test cases
that span the majority of observation types likely to be undertaken
with \scuba, with an emphasis on the mitigation of divergence
problems, and characterizing the output maps: (i) Uranus, a bright,
compact source (Section~\ref{sec:point}); (ii) the Lockman Hole, a
blind survey of faint point-like source (Section~\ref{sec:cosmo}); and
(iii) the star-forming region M17, including bright, extended emission
(Section~\ref{sec:extended}). All of the data analysed in this paper
are publicly available through the CADC \scuba\ raw-data queries page
for the dates and observation numbers given in the
text\footnote{\url{http://www.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/jcmt/search/scuba2}}.
All of the analysis was performed using the Starlink \textsc{kapuahi}
release from 2012.

@@

As described in the introduction, the approach taken by SMURF to
reducing \scuba\ data is to model (and remove) predominantly
low-frequency noise sources that are correlated amongst detectors, and
to iterate this process along with estimates of the map. Significant
experimentation with different models for noise sources during \scuba\
commissioning required a highly flexible software framework, and a
configurable user interface. To achieve this goal, while minimizing
development time, it was decided to build SMURF as a Starlink package
\citep{2009ASPC..411..418J}, which provides access to a large suite of
libraries (including a commanding and messaging interface, astrometric
coordinate conversions and generation of standard WCS information,
file formats etc.). Furthermore, Starlink\footnote{\url{http://www.starlink.ac.uk}} is
open-source\footnote{\url{https://github.com/Starlink}},
and already used extensively at the Joint Astronomy Centre (host of the JCMT) for
many other systems \citep{jenness2011}, which helps with interoperability. Though
originally written in FORTRAN, many of the core Starlink libraries are
now ported to native C, or at least have a C interface. It was
therefore decided to develop SMURF in C as well (rather than C++, for
example, which would have required adding further dependencies to
Starlink), although we have taken an object-oriented approach. For
example, all of the data for a given noise model is encapsulated in a
C structure, and there is a standard interface for all functions that
handle models (member data and functions for the class,
respectively). In this way it is easy to extend SMURF with new
models. Parallelisation is incorporated in the most time-consuming
low-level routines using threads (e.g., performing Fast Fourier
Transforms, re-gridding the data), usually handling either independent
blocks of bolometers, or blocks of time, in each thread, depending on
the nature of the calculation. \textbf{... something about scalability
...}
